draft 1

### 2016-06-29 finalwriteup/draft.rmd
finished rejected jobs analysis

**Highest Profit Curves**

@@

**Comparison of All Profit Curves**

@@

**Comparison of Confusion Matrix Statistics**

@@

All individual original methods used substantially lower TP rates at their optimal thresholds, while their TN rates were spread out. This means, as the trialled threshold values moved, and the TP rate improved, the increasing lost profits from FP's was too great to justify a TP rate higher than 4.6%. Blending the individual methods has made use of their individual strengths and seems to have clustered the probabilities of the actual loss making jobs more accurately towards 1, and the profitable jobs towards 0. This allowed for higher TP rates with lower FP rates at the blended methods' optimal thresholds.

**Review of Jobs Rejected by Simple Logistic Regression**

All below analyses for results of a typical Simple Blended Logistic Regression Model

* ratio of profitable to unprofitable

```{r}
reject.plot <- readRDS('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/reject_plot.rds')

@@

* size of projects in profitable and unprofitable divisions

```{r}
ggplot(reject.plot %>% filter(compare == "reject"), 
             aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
        geom_bar(stat = "identity") + 
        theme(legend.position="none") +
        scale_fill_manual(values = rep(c('dodgerblue4', 'darkolivegreen1'), 59)) + 
        scale_y_continuous(labels = scales::dollar) +
        labs(x = "Project Performance", 
             y = "Cumulative Profits/Losses",
             title = "Magnitude of Profits/Losses from Rejected Projects") +
        geom_text(data = NULL, x = .5, y = 720000, size = 3, 
                  colour = "gray60", hjust = 0,
                  label = "Alternate Colours \nIndicate Different \nProjects")

@@

* some plots of variables that vary quite a bit
    * pc.pro

@@

    * timespan kind of interesting: no 1d - 3wk or 3wk to 2.5 m

@@

```
    * return per dollar quite nice

@@

Not in a plot, but state:
    * majority position: mid tech and senior tech disproportionately high
    * JD.Second: 16 categories originally.
        * subdivision and sewerage disproportionately high

**Alternate Interpretations of Blended Model Results**

@@

@@

>>>>

### Highest Profit Curves

@@

### Comparison of All Profit Curves

@@

### Comparison of Confusion Matrix Statistics

@@

All individual original methods used substantially lower TP rates at their optimal thresholds, while their TN rates were spread out. This means, as the trialled threshold values moved, and the TP rate improved, the increasing lost profits from FP's was too great to justify a TP rate higher than 4.6%. Blending the individual methods has made use of their unique strengths and seems to have clustered the probabilities of the actual loss-making jobs more accurately towards 1, and the profitable jobs towards 0. This allowed for higher TP rates with lower FP rates at the blended methods' optimal thresholds.

### Review of Jobs Rejected by Simple Logistic Regression

The profit curves and TP/FP rates give a good indication of overall performance of the methods, however an analysis of the characteristics of the predicted loss-making projects can provide more detailed insight.

All below plots and comments summarise results from one Simple Blended Logistic Regression Model. The results from this model were chosen because it represented the mean for improvement in profits output by this blending method. The plot below illustrates how the ratio of profitable to unprofitable jobs compare in the full project population vs. the rejected projects.

```{r}
reject.plot <- readRDS('report_data/reject_plot.rds')

@@

Originally, 25% of projects were loss-making. The model rejects a population of projects which are over 60% loss-making. This represents a significant swing in proportions, but is not perfect. The class imbalance towards profitable jobs in the initial dataset may have weakened the model's ability to classify losses. @Saradhi2011 took advantage of SVM's ability to adjust class penalties to address class imbalance, however SVM's were avoided because of their black box predictions in this work and the remaining methods cannot accommodate class penalties.

The plot below illustrates the monetary size of profits and losses from rejected projects. The different coloured bands indicate different projects and it can be observed that the magnitude of rejected profits and losses vary greatly. They range from profits/losses of just \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% tail(1) %>% round(0)` to losses of \$`r reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))) %>% .$balance.mlsto %>% head(1) %>% round(-3)`.

```{r}
ggplot(reject.plot %>% filter(compare == "reject") %>% arrange(desc(abs(balance.mlsto))), 
             aes(x = f.rpdol, y = abs(balance.mlsto), fill = mlsto)) +
        geom_bar(stat = "identity") + 
        theme(legend.position="none") +
        scale_fill_manual(values = rep(c('dodgerblue4', 'darkolivegreen1'), 59)) + 
        scale_y_continuous(labels = scales::dollar) +
        labs(x = "Project Performance", 
             y = "Cumulative Profits/Losses",
             title = "Magnitude of Profits/Losses from Rejected Projects") +
        geom_text(data = NULL, x = .5, y = 720000, size = 3, 
                  colour = "gray60", hjust = 0,
                  label = "Alternate Colours \nIndicate Different \nProjects")

@@

The variable 'percentage of hours completed by a professional level employee' had an interesting contrast in the statistical distribution of all projects vs. the rejected projects (refer plot below).

@@

The algorithm tended to accept projects which were completed 100% by a professional level employee and rejected more projects with 0-60% of hours by a professional employee. This could either indicate that professional level employees complete projects more efficiently or that the types of projects that do not require technical employees are more profitable.

The distribution of timespans of rejected projects differed significantly to the timespan distribution of all projects. The chart below shows that proportionally very few short jobs were rejectd, i.e. projects that lasted between 1 day and 2.5 months. These projects are likely small and may be easier to estimate fixed fees. A much higher relative proportion of projects lasting 1.5 - 3 years were rejected. They may face an opposite challenge to the small projects. On the other hand, large projects can provide long term security for a small business and it is unlikely the company would want to reject them. It may make more sense to adjust the contract structure of large projects that have been highlighted as a risk. Alternate uses of the algorithm results are discussed in the next section.

@@

The statistical distribution of return per dollar in the rejected projects are skewed towards negative returns. This means the algorithm has identified a proportionally much higher percentage of loss-making projects which is not surprising given previous results.

@@

There were a couple more revealing differences between the full set of projects and the rejected projects. The first was that rejected projects had disproportionately more projects with mid-level and senior-level technical employees as the main worker on a project. The second was that a comparatively high percentage of rejected projects were subdivisions, sewerage, or civil building works.

The detailed analysis of rejected project characteristics gave insight into the characteristics of projects that are likely to be loss-making as well as the accuracy of the algorithm. While it is interesting to observe characeristics of rejected projects, a complex combination of all the features contribute to the final probability of loss prediction.

### Alternate Interpretations of Blended Model Results

@@

### User Interface and Engagement

* Highlighted as a major problem in previosu cost estimation studies
    * need to (1) communicate results usefully (2) trust
* User actions will vary, options listed in previous section: reject project, adjust contract structure
    * Communicate confidence
    * future work: probability: bayesian?
    * simple yes/no, easy, or on a scale of probability 0.6 to 1, but output is not statistically driven. score driven by machine learning algorithms.
    * future work: which is preferable user chooses

* Shiny App: engage user trust by two strategies
    * nearest neighbours
    * model output

@@

**updating the model!**
A lot of data cleaning in past data
need to build a pipeline of cleaning - only 11 variables now, not too hard
* re run algorithm monthly? does not take long, maybe 1 minute.
* repeat validation every few months
